---
title: "Covid_procurements"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(stringr)
library(ggstatsplot)
library(splusTimeDate)
```

## Merge & clean Files

The files results_all_records_x were scrapped from the Mission Mode Portal over two weeks. 
These files contain "Sno", "e_published date", "Award of contract date" "Title and Description" and "State". We will first merge the four files that were scrapped and clean them

```{r Merge, echo=FALSE}
## read the csvs as a tibble for easier tidyverse stuff (especially string replace which requires it)
results_1 <- as.tibble(read.csv("results_all_records_table.csv", row.names = NULL,
                                stringsAsFactors = FALSE))
results_1a <- as.tibble(read.csv("results_all_records_table_1.csv", row.names =NULL,
                                stringsAsFactors = FALSE))

results_2 <- as.tibble(read.csv("results_all_records_table_2.csv", row.names = NULL, 
                                stringsAsFactors = FALSE))
results_3 <- as.tibble(read.csv("results_all_records_table_3.csv",row.names = NULL, 
                                stringsAsFactors = FALSE))
results_4 <- as.tibble(read.csv("results_all_records_table_4.csv",row.names = NULL, 
                                stringsAsFactors = FALSE))
results_extra <- as.tibble(read.csv("results_all_records_table_extra.csv", 
                           row.names = NULL, stringsAsFactors = FALSE))

results_covid <- as_tibble(read.csv("Result_state_tenders.csv", 
                                    row.names = NULL, stringsAsFactors = FALSE))

central_covid <- as_tibble(read.csv("Central_tenders.csv", 
                                    row.names = NULL, stringsAsFactors = FALSE))

## merge the four files with rbind
results_total_merge<- rbind(results_1,results_1a,results_2,results_3,results_4, results_extra)

## clean the merged file of the b' that .' that is added as a part of the python scrapper

results_total_merge_clean<- results_total_merge %>%
    mutate_all(funs(str_replace(., "b'", "")))

results_total_merge_clean <- results_total_merge_clean %>% 
  mutate_all(funs(str_replace(., "'", "")))

results_total_merge_clean <- results_total_merge_clean %>% 
  mutate_all(funs(str_replace(., "\\.", "")))

## rename the columns
results_total_merge_clean <- results_total_merge_clean %>% 
  rename(
    Sno= row.names,
    award_date = S.No.,
    Published_date = AOC.Date,
    Title_description = e.Published.Date, 
    State_name = Title.and.Ref.No..Tender.Id, 
    Tender_ID = State.Name
    )

results_covid <- results_covid %>% rename (
    Sno= S.No.,
    award_date = AOC.Date,
    Published_date = e.Published.Date,
    Title_description = Title.and.Ref.No..Tender.Id, 
    State_name = State.Name, 
    worktype = Type.of.work, 
    Purchase_classification = What.is.it, 
    Number_bids = Number.of.bids, 
    Bid_winner = Bid.winner)

central_covid <- central_covid %>% rename (
    Sno= Sno,
    award_date = AOC.date,
    Published_date = Published.date,
    Title_description = Tender.description, 
    State_name = Org.name, 
    worktype = Type.of.Work, 
    Purchase_classification = Purchase.type, 
    Contract_value = Contract.value,
    Bid_winner = Supplier,
    Number_bids = Number.of.bids)

central_covid$Purchase_classification <- toupper(central_covid$Purchase_classification)

results_covid$Purchase_classification <- toupper(results_covid$Purchase_classification)

## Make dates into date format
results_total_merge_clean$award_date <- as.Date(results_total_merge_clean$award_date, 
                                                format="%d-%b-%Y")

results_total_merge_clean$Published_date <- as.Date(results_total_merge_clean$Published_date, 
                                                    format="%d-%b-%Y")


results_covid$award_date <- as.Date(results_covid$award_date, 
                                                format="%d-%b-%Y")

results_covid$Published_date <- as.Date(results_covid$Published_date, 
                                                format="%d-%b-%Y")

central_covid$award_date <- as.Date(central_covid$award_date, format="%d-%b-%Y")
central_covid$Published_date <- as.Date(central_covid$Published_date, format="%d-%b-%Y")
 
##remove duplicates
results_no_duplicates <- results_total_merge_clean [!duplicated(results_total_merge_clean[c(4,6)]),]

##Clean the tender description into two parts - tender name and ref number and org id
results_covid <- results_covid %>% separate (Title_description, into = c("Title_name","Ref_ID"), sep = "\\/", extra = "merge")

results_covid <- results_covid %>% 
   separate(Ref_ID, into = c("Ref", "Tender ID"), sep="/(?=[^/]+$)")

results_covid<- subset(results_covid, select = -c(13:22))

```

## Initial look

We can have an initial look at the data that we have cleaned so far. For example we can see which states have the most number of tenders and what the different durations for the tendering process are

```{r initial, echo=FALSE}
## this will produce the top ten states that have most number of procurements on mmp

tender_number <- results_no_duplicates %>% group_by(State_name) %>% tally()
head(tender_number[order(tender_number$n, decreasing = TRUE),])

results_covid_state<- results_covid%>% group_by(State_name) %>% tally()
head(results_covid_state[order(results_covid_state$n, decreasing = TRUE),])

results_covid_cost<- results_covid%>% group_by(State_name) %>% summarise (total = sum(as.numeric(Amount), na.rm = TRUE))
head(results_covid_cost[order(results_covid_cost$total, decreasing = TRUE),])

## Summary: We have seen the top 10 covid procuring states by number of procurements and amounts. Maharashtra has spent half as much on 127 procurements as Madhya Pradesh did on 9. However, many of the awards information for procurements have not been inputted correctly. Some of them have mentioned a single line items price but not the price for supplying that item for the year. This could bring in some degree of error

## Here we are creating a simple bar plot with counts for each state
results_no_duplicates %>% ggplot(aes(x = State_name))+ 
  geom_bar ()+
  theme(axis.text.x = element_text(size = 6, angle = 90))

plot1 <- results_covid %>% ggplot(aes(x = State_name))+ 
  geom_bar ()+
  theme(axis.text.x = element_text(size = 6, angle = 90))
plot1

#Here we have calculated the number of days between the published and awarded date
results_no_duplicates$number_days<-
  as.numeric(difftime (results_no_duplicates$award_date, 
                       results_no_duplicates$Published_date,units = c("days")))


results_covid$number_days <- as.numeric(difftime (results_covid$award_date, 
                       results_covid$Published_date,units = c("days")))

## we are now trying to create a stacked box plot to see the difference in days taken
results_no_duplicates %>% ggplot(aes(x = State_name, fill = number_days))+ 
  geom_bar ()+
  theme(axis.text.x = element_text(size = 6, angle = 90))

plot2 <- results_covid %>% ggplot(aes(x=State_name, fill = number_days))+  
  geom_bar ()+
  theme(axis.text.x = element_text(size = 6, angle = 90))

plot2

##Summary: Not a great way to visualize the variation in days or number of bidders. so lets make a box plot instead
## Since the variation in the other plot is so big, lets make a box plot and see how it looks

results_no_duplicates %>% ggplot(aes(x=State_name, y=number_days)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))

plot3 <- results_covid %>% ggplot(aes(x=State_name, y=number_days)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))


plot4 <- results_covid %>% ggplot(aes(x=State_name, y=Number_bids)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))

## Summary: Here it is clearly seen that while some states have number of days for procurement as less than 100, with chandigarh and delhi around 25 days. Other states such as Assam, Haryana have mean procurements days stretching up to a 150. While most states have a similar number of bidders per tender hovering around 2. There are some states that perform slightly better like punjab and delhi which have about 7-8 bids per tender. 

##We will now do similar calculations with type of tender
plot5 <- results_covid %>% ggplot(aes(x=worktype, y=number_days)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))
plot6 <- results_covid %>% ggplot(aes(x=worktype, y=Number_bids)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))

## Summary: Interestingly tenders around services seems to be concluded the fastest at a mean 50 days while works are slightly more and goods are highest. Though there are more exceptions in works. This may not be a significant difference. (could check). Services and goods also seem to have the most number of bidders followed by works.
results_covid$Purchase_classification <- gsub('COVID ALLIED WORK', 'COVID ALLIED',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID PREVENTIOB', 'COVID PREVENTION',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID MEDICATION', 'COVID TREATMENT',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID MANAGEMNT', 'COVID INFORMATION',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID HCW - DIET', 'COVID HCW',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID WARD - PREGNANT WOMEN', 'COVID WARD -WOMEN',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID WARD -WOMEN', 'COVID WARD - WOMEN',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID TESTING', 'COVID DETECTION',results_covid$Purchase_classification)

results_covid$Purchase_classification <- gsub('COVID ROAD', 'COVID ALLIED',results_covid$Purchase_classification)


plot7 <- results_covid %>% ggplot(aes(x=Purchase_classification, y=Number_bids)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))
plot8 <- results_covid %>% ggplot(aes(x=Purchase_classification, y=number_days)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))

## When we look at the classification of covid tasks, covid allied work took the most number of days for contracts followed by covid informational tasks. Tenders involving covid body removal and diet and health care workers took the least amount of time for award of contract. Covid diet and covid detection had more number of bidders per tender.

## Now I will cut the year up into three parts 
results_covid <- 
  results_covid %>% 
  separate(col = Published_date, into = c('yr', 'mnth', 'dy'), sep = '-', convert = TRUE, remove = FALSE)


  # make quarters
results_covid <- 
  results_covid %>% 
  mutate(
    quarter = case_when(
      mnth >= 6 & mnth <= 8 ~ 'Q1'
      , mnth >= 9 & mnth <= 10 ~ 'Q2'
      , mnth >=  11 & mnth <= 1 ~ 'Q3'))

results_covid_grp <-  results_covid %>% group_by(mnth, Purchase_classification) %>% summarise (total = sum(as.numeric(Amount), na.rm = TRUE))

plot9 <- results_covid_grp %>% ggplot(aes(x=mnth, y=total, fill = Purchase_classification)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(size = 6, angle = 90))

## There seems to be quite a bit of variation, so i will be identifying the outliers to eliminate
Q <- quantile(results_no_duplicates$number_days, probs=c(.5, .9), na.rm = FALSE)

iqr <- IQR(results_no_duplicates$number_days)

up <-  Q[2]+1.5*iqr
down <- Q[1] - 1.5*iqr

not_eliminated<- subset(results_no_duplicates, 
                        results_no_duplicates$number_days > (Q[1] - 1.5*iqr) 
                        & results_no_duplicates$number_days < (Q[2]+1.5*iqr))

## I will then plot the new dataset to see how it has changed
##outliers <- ggbetweenstats(data = results_no_duplicates, x = State_name, y=number_days, outlier.tagging = TRUE, na.rm = TRUE) 

## I am seeing whats there in the eliminated (outliers)
eliminated <- subset(results_no_duplicates, 
                     !(results_no_duplicates$number_days > 
                         (Q[1] - 1.5*iqr) & results_no_duplicates$number_days < 
                         (Q[2]+1.5*iqr)))

eliminated %>% ggplot(aes(x=State_name, y=number_days)) + geom_boxplot() + 
  theme(axis.text.x = element_text(size = 6, angle = 90))

##I have decided to use the non-outliered dataset since, the data in the eliminated set seems important

## I am creating bins for the length of the different procurement periods according to state

results_no_duplicates$day_category <- cut(results_no_duplicates$number_days, 
                                          breaks = c(0, 30, 60, Inf))

## I am plotting this data using a stacked bar graph

results_no_duplicates %>% ggplot(aes(x=State_name, fill = day_category)) + 
  geom_bar() +theme(axis.text.x = element_text(size = 6, angle = 90))

```

## Next I will create smaller datasets for specific usecases. The following is a table for words related to health and covid. Isolating all procurements associated with the keywords will give us a subset of health and covid related procurements. (Note: while health is a general enough term, I have taken caution not to include sanitation under this. Health in this case will only comprise of acute/chronic hospital related procurements not general sanitation or food/nutrition)

```{r pattern, echo=FALSE}
patterns <-data.frame(word = c("Medical", "Hospital", "PHC", "CHC", "Health", "covid", "COVID", "medicine", "drugs", "liquid", "ambulance",  "microscope", "chemistry", "radiograph", "biomedical", "laboratory", "dispensary", "Anganwadi","DHS","subcentre", "ANM", "sub centre", "HWC", "RBSK", "JSSK", "KASP","crematorium","mortem","paediatric","^MCH","Smashanbhumi","Shmashanbhumi","smashan bhumi","HFW","ICDS","RT-PCR","PPE","corona","anganwari","NHM","Advia","Oxygen","^HCG","AWC$","mask"))

#pattern_match <- data.frame(word = c("Medical","Hospital","PHC","CHC","Health","covid","medicine","drugs","liquid", "ambulance","microscope","chemistry", "radiograph","biomedical","lab","laboratory","dispensary","Anganwadi","DHS","subcentre", "ANM", "sub centre", "HWC", "RBSK", "JSSK", "KASP", "crematorium", "DNASTAR","mortem", "paediatric", "HWC", "MCH", "Smashanbhumi", "Shmashanbhumi", "smashan bhumi", "OPD", "HFW", "ICDS","RT-PCR","corona","anganwari","NHM", "TCL","Advia","Oxygen","HCG","AWC","DGPCS"), matching = c("1l", "2", "3", "4","5", "6", "7", "8", "9", "10","11","12","13","14","15","16","17","18","19","20","21","22","23","24","25","26","27","28","29","30","21","32","33","34","35","36", "37", "38","39","40","41", "42", "43", "44", "45", "46", "47", "48"))

  
newtable <- results_no_duplicates %>% 
 filter_all(any_vars(str_detect(.,paste(patterns$word, collapse = "|"))))

newtable <- newtable %>% 
  filter(!grepl("Vet",Title_description))

#supply, construction,maintanence, repair, up-gradation, upgradation, Improvement, Installation, Labour, facility, printing, work, outsource, manpower, const, Reconstruction, handling, service, providing, purchase, strenghthening, painting, road?,replacing, training,transport, Running, nirman, establishment, renovation, 

## Different ways to add keywords. The non-commented out one works best (need to figure out how to add more than 1 key word)
#newtable <- newtable%>% 
#mutate(newcol = str_replace(Title_description, pattern_match$word, pattern_match$matching, vectorize_all = FALSE))

#for(i in seq_len(nrow(pattern_match))) {
 #want <- grepl(pattern_match[i, "word"], newtable[,"Title_description"],ignore.case = T)
 #newtable[want, "matching"] <- pattern_match[i, "matching"]
#} 
#results_no_duplicates$newcol <- results_no_duplicates %>% 
#paste(any_vars(str_detect(.,paste(patterns$word, collapse = "|"))))

#newtable$keywords <- str_extract(newtable$Title_description, str_flatten(patterns$word, "|"))

toMatch<-c("health","hospital", "PHC", "CHC", "Hospital", "subcentre", "sub centre", "^SC$", "HWC", 
           "sub centre", "ambulance", "medical", "medicine", "Sub Centre","diet", "^DHS", "^DMC", "Health", "Medical", "Suswasthya", "DIETARY", "HEALTH","\\bKASP\\b","\\bDHS\\b", "Food", "surgical", "drugs, Anesthesia", "ICU", "DIALYSIS", "Ayushman", "AYUSHMAN", "P.H.C","MEDICAL", "Sub Centers", "SUB CENTER", "HOSPITALS", "Medicine", "DOCTOR", "HOSPITAL", "Mammography", "MEDICINE", "Diet", "Med", "Tepania", "Canteen", "canteen")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description),"keywords"]<-"Health Care Centres"

toMatch<-c("chemistry", "College","Department", "//bMCH//b", "laboratary", "COLLEGE", "laboratory")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description ),"keywords"]<-"Medical College"

toMatch<-c("\\bANM\\b","^anganwari", "Anganwadi","\\bAWC\\b", "^paediatric", "MOTHER", "CHILD", "ANGANWARI", "ASHA", "\\bICDS\\b","\\bJSSK\\b","\\bRBSK\\b", "AGANWADI")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description ),"keywords"]<-"Paediatric"

toMatch<-c("Water", "WATER","^air$", "^Air$", "water", "TUBEWELL", "HANDPUMP", "Jal Jeevan")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description ),"keywords"]<-"Water & Air Quality"

toMatch<-c("crematorium", "Smashanbhumi", "Shmashanbhumi","^smashan bhumi","^dead", "mortem", "Mortuary")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description ),"keywords"]<-"Death"

toMatch<-c("^covid", "^corona","^cylinder", "^oxygen","^RT-PCR", "^liquid","\\bPPE\\b","^mask", "^drugs","^Advia", "Sanitizer", "DISINFECTANT", "masks", "mask", "Hypochlorite", "COVID", "Covid", "Corona", "CORONA", "Isolation", "isolation", "quaranti", "Quaranti", "Quarenti", "quarenti", "Oxygen", "OXYGEN", "covid-19", "corona", "RT-PCR", "RTPCR", "VTM", "oxygen")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description),"keywords"]<-"COVID"

toMatch<-c("Stationary","OUTSOURCE", "FURNITURE", "Computer", "Laptop", "Photocopy", "PRINTERS", "vehicle", "Vehicle", "House Keeping", "Recruitment", "Printing")
#newtable$keywords<-as.character(newtable$keywords)
newtable[grepl(paste(toMatch,collapse="|"),newtable$Title_description),"keywords"]<- "Allied"

Q <- quantile(newtable$number_days, probs=c(.5, .9), na.rm = FALSE)

iqr <- IQR(newtable$number_days)

up <-  Q[2]+1.5*iqr
down <- Q[1] - 1.5*iqr

newtable<- subset(newtable, 
           newtable$number_days > (Q[1] - 1.5*iqr) 
           & newtable$number_days < (Q[2]+1.5*iqr))

state_tenders<- newtable %>% group_by(State_name) %>% tally()
head(state_tenders[order(state_tenders$n, decreasing = TRUE),])

newtable %>% ggplot(aes(x = State_name))+ 
  geom_bar ()+
  theme(axis.text.x = element_text(size = 6, angle = 90))

newtable %>% ggplot(aes(x=State_name, y=number_days)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size = 6, angle = 90))

newtable %>% ggplot(aes(x=State_name, fill = day_category)) + 
  geom_bar() +theme(axis.text.x = element_text(size = 6, angle = 90))


newtable$Month <- format(as.Date(newtable$Published_date, "%Y/%m/%d"), "%m")

newtable %>% ggplot(aes(x=Month, fill = day_category)) + 
  geom_bar() +theme(axis.text.x = element_text(size = 6, angle = 90))

newtable_sep <-  newtable %>% group_by(State_name, keywords) %>% tally()
newtable_mnth <-  newtable %>% group_by(Month, keywords) %>% tally()
newtable_number <- newtable %>% group_by(number_days, keywords) %>% tally()

plot_newtable <- newtable_sep %>% ggplot(aes(x=State_name, y=n, fill = keywords)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(size = 6, angle = 90))

plot_newtable1 <- newtable_mnth %>% ggplot(aes(x=Month, y=n, fill = keywords)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(size = 6, angle = 90))

plot_newtable2 <- newtable_number %>% ggplot(aes(x=number_days, y=n, fill = keywords)) + 
  geom_bar(stat = "identity") + theme(axis.text.x = element_text(size = 6, angle = 90))


##next need to remove hwc if in haryana

write.csv2(newtable, "newtable.csv")
```

